---
title: "Phage and nitrogen experiment"
author: "John McMullen"
date: "11/29/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

Late growth experiment to look at the interaction between rhizobia, nitrogen, and lytic phages
Note: the aposymbiotic plants were contaminated with rhizobia (nodules present).

#Libraries
```{r}
library(car) #Anova
library(tidyverse) #data wrangling and plotting
library(cowplot) #plotting
library(emmeans) #post hoc testing
library(multcomp) #letter ranking
library(multcompView) #letter ranking

library(vegan) #PCA
library(corrplot) #corrplot.mixed function
library(PerformanceAnalytics) #chart.Correlation
library(psych) #Bartlett's test and KMO
```

#Data
```{r}
dat = read.csv(file = "late-growth_phage_nitrogen_461_262.csv")

dat = dat %>%
  mutate(micro = ifelse(micro %in% "461", "HQ",
                        ifelse(micro %in% "262", "LQ","Apo")))

dat$micro = as.factor(dat$micro)
dat$N = as.factor(dat$N)
dat$P = as.factor(dat$P)
```


Cohen's d is the mean - mean/pooled sd
for ci, take 1.645 * sqrt((n+n/n*n)+(d^2/2*(n+n)))

#stolon - effect size
```{r}
stol_mean = dat %>%
  dplyr::select(N,P,micro,stolon) %>%
  filter(!micro %in% "Apo") %>%
  group_by(N,P,micro) %>%
  summarise(mean = mean(stolon),count = length(stolon)) %>%
  pivot_wider(values_from=mean,names_from=P)

stol_sd = dat %>%
  dplyr::select(N,P,micro,stolon) %>%
  filter(!micro %in% "Apo") %>%
  group_by(N,micro) %>%
  summarise(sd = sd(stolon))

stol_dat = stol_mean %>%
  left_join(stol_sd,by = c("N"="N","micro"="micro")) %>%
  group_by(micro,N) %>%
  mutate(d = ((`1` - `0`)/sd)) %>%
  mutate(ci = 1.645* sqrt((count + count) / (count * count) + d^2 / (2 * (count + count)))) %>%
  mutate(fit_axis = "Stolon")
```

#flower - effect size
```{r}
flow_mean = dat %>%
  dplyr::select(N,P,micro,flower) %>%
  filter(!micro %in% "Apo") %>%
  group_by(N,P,micro) %>%
  summarise(mean = mean(flower),count = length(flower)) %>%
  pivot_wider(values_from=mean,names_from=P)

flow_sd = dat %>%
  dplyr::select(N,P,micro,flower) %>%
  filter(!micro %in% "Apo") %>%
  group_by(N,micro) %>%
  summarise(sd = sd(flower))

flow_dat = flow_mean %>%
  left_join(flow_sd,by = c("N"="N","micro"="micro")) %>%
  group_by(micro,N) %>%
  mutate(d = ((`1` - `0`)/sd)) %>%
  mutate(ci = 1.645* sqrt((count + count) / (count * count) + d^2 / (2 * (count + count)))) %>%
  mutate(fit_axis = "Flower")
```

#flower - effect size
```{r}
mass_mean = dat %>%
  dplyr::select(N,P,micro,plant) %>%
  filter(!micro %in% "Apo") %>%
  group_by(N,P,micro) %>%
  summarise(mean = mean(plant),count = length(plant)) %>%
  pivot_wider(values_from=mean,names_from=P)

mass_sd = dat %>%
  dplyr::select(N,P,micro,plant) %>%
  filter(!micro %in% "Apo") %>%
  group_by(N,micro) %>%
  summarise(sd = sd(plant))

mass_dat = mass_mean %>%
  left_join(mass_sd,by = c("N"="N","micro"="micro")) %>%
  group_by(micro,N) %>%
  mutate(d = ((`1` - `0`)/sd)) %>%
  mutate(ci = 1.645* sqrt((count + count) / (count * count) + d^2 / (2 * (count + count)))) %>%
  mutate(fit_axis = "Biomass")
```

#effect size plot
```{r}
effect_dat = stol_dat %>%
  ungroup() %>%
  add_row(flow_dat %>% ungroup()) %>%
  add_row(mass_dat %>% ungroup())

effect_dat$micro = factor(effect_dat$micro,levels=c("LQ","HQ"))
effect_dat$fit_axis = factor(effect_dat$fit_axis,levels=c("Stolon","Flower","Biomass"))

ggplot(data = effect_dat,
       aes(x = d, y = fit_axis, color = N)) +
  geom_errorbar(aes(xmin = d-ci, xmax = d+ci),width=0.1,position = position_dodge(width = 0.8)) +
  geom_vline(xintercept = 0,linetype="dashed") +
  geom_point(size = 5,position = position_dodge(width = 0.8)) +
  scale_color_manual(values=c("black","darkgray")) +
  scale_x_continuous(limits = c(-3,2),breaks = seq(-3,2,by=1)) +
  xlab(expression("Cohen's D \u00B1 CI")) +
  ylab("Fitness components") +
  facet_grid(.~micro) +
  theme_bw(base_size = 20) +
  theme(panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black",fill=NA),
        panel.background = element_rect(fill="white"),
        axis.text.x=element_text(size=rel(0.9)),
        axis.text.y=element_text(size=rel(0.9)),
        axis.title.x =element_text(size=rel(0.9)),
        axis.title.y = element_text(size=rel(0.9)))
```

##########################
Supplemental

#Plot stolon
```{r}
stolon_apo_mean = dat %>%
  filter(micro %in% "Apo") %>%
  group_by(N) %>%
  summarise(mean = geometric.mean(stolon))

#removed nonsignificant mixed effects to simplify
stol.mdl = glm(stolon ~ micro * N * P,family = poisson(link = "log"),data = dat %>% filter(!micro %in% "Apo"))

AIC(stol.mdl)

summary(stol.mdl)

#assess normality
stol.res = residuals(stol.mdl)
summary(stol.res)
qqPlot(stol.res)
hist(stol.res)
boxplot(stol.res)

#assess homoscedasticity
plot(fitted(stol.mdl),stol.res);
abline(h=0,lty=2);
lines(smooth.spline(fitted(stol.mdl),stol.res));

#ANOVA test table
Anova(stol.mdl,type="II")

#test for post hos testing
stol.emmeans = emmeans(stol.mdl, ~ P | N*micro,type="response",level=0.9)
pairs(stol.emmeans)

#plot
stolon = ggplot(data = as.data.frame(stol.emmeans),
       aes(x = P, y = rate, color = N, shape = P, group = N, fill = P)) +
  geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL),width=0.1,position = position_dodge(width = 0.5)) +
  geom_line(position = position_dodge(width = 0.5))+
  geom_hline(yintercept = 4.9,linetype="dashed") +
  geom_hline(yintercept = 9.5,linetype="dashed",color="darkgray") +
  geom_point(size = 5,position = position_dodge(width = 0.5)) +
  scale_color_manual(values=c("black","darkgrey")) +
  scale_shape_manual(values=c(21,19)) +
  scale_fill_manual(values = c("white","white")) +
  scale_y_continuous(limits = c(0,15),breaks = seq(0,15,by=5)) +
  facet_grid(.~factor(micro,levels = c("Apo","LQ","HQ"))) +
  ylab(expression(atop("Stolon count","mean \u00B1 CI (n = 5)"))) +
  xlab("Phage addition") +
  theme_bw(base_size = 20) +
  theme(panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black",fill=NA),
        panel.background = element_rect(fill="white"),
        axis.text.x=element_text(size=rel(0.9)),
        axis.text.y=element_text(size=rel(0.9)),
        axis.title.x =element_text(size=rel(0.9)),
        axis.title.y = element_text(size=rel(0.9)),
        legend.position = "none")
```

#Plot flower
```{r}
flower_apo_mean = dat %>%
  filter(micro %in% "Apo") %>%
  mutate(flower = ifelse(flower == 0, 1, flower)) %>%
  group_by(N) %>%
  summarise(mean = geometric.mean(flower))

#removed nonsignificant mixed effects to simplify
flow.mdl = glm(flower ~ micro * N * P,family = quasipoisson(link = "log"),data = dat %>% filter(!micro %in% "Apo"))

AIC(flow.mdl)

summary(flow.mdl)

#assess normality
flow.res = residuals(flow.mdl)
summary(flow.res)
qqPlot(flow.res)
hist(flow.res)
boxplot(flow.res)

#assess homoscedasticity
plot(fitted(flow.mdl),flow.res);
abline(h=0,lty=2);
lines(smooth.spline(fitted(flow.mdl),flow.res));

#ANOVA test table
Anova(flow.mdl,type="II")

#test for post hos testing
flow.emmeans = emmeans(flow.mdl, ~ P | N*micro,type="response",level=0.9)
pairs(flow.emmeans)

#plot
flower = ggplot(data = as.data.frame(flow.emmeans),
       aes(x = P, y = rate, color = N, shape = P, group = N, fill = P)) +
  geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL),width=0.1,position = position_dodge(width = 0.5)) +
  geom_line(position = position_dodge(width = 0.5))+
  geom_hline(yintercept = 6.84,linetype="dashed") +
  geom_hline(yintercept = 8.07,linetype="dashed",color="darkgray") +
  geom_point(size = 5,position = position_dodge(width = 0.5)) +
  scale_color_manual(values=c("black","darkgrey")) +
  scale_shape_manual(values=c(21,19)) +
  scale_fill_manual(values = c("white","white")) +
  scale_y_continuous(limits = c(0,50),breaks = seq(0,50,by=10)) +
  facet_grid(.~factor(micro,levels = c("Apo","LQ","HQ"))) +
  ylab(expression(atop("Flower count","mean \u00B1 CI (n = 5)"))) +
  xlab("Phage addition") +
  theme_bw(base_size = 20) +
  theme(panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black",fill=NA),
        panel.background = element_rect(fill="white"),
        axis.text.x=element_blank(),
        axis.text.y=element_text(size=rel(0.9)),
        axis.title.x =element_blank(),
        axis.title.y = element_text(size=rel(0.9)),
        legend.position = "none")
```

#Plot biomass
```{r}
biomass_apo_mean = dat %>%
  filter(micro %in% "Apo") %>%
  group_by(N) %>%
  summarise(mean = mean(plant))

#removed nonsignificant mixed effects to simplify
biomass.mdl = lm(plant ~ micro * N * P,data = dat %>% filter(!micro %in% "Apo"))

AIC(biomass.mdl)

summary(biomass.mdl)

#assess normality
biomass.res = residuals(biomass.mdl)
summary(biomass.res)
qqPlot(biomass.res)
hist(biomass.res)
boxplot(biomass.res)

#assess homoscedasticity
plot(fitted(biomass.mdl),biomass.res);
abline(h=0,lty=2);
lines(smooth.spline(fitted(biomass.mdl),biomass.res));

#ANOVA test table
Anova(biomass.mdl,type="II")

#test for post hos testing
biomass.emmeans = emmeans(biomass.mdl, ~ P | N*micro,type="response",level=0.9)
pairs(biomass.emmeans)

#plot
mass = ggplot(data = as.data.frame(biomass.emmeans),
       aes(x = P, y = emmean, color = N, shape = P, group = N, fill = P)) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),width=0.1,position = position_dodge(width = 0.5)) +
  geom_line(position = position_dodge(width = 0.5))+
  geom_hline(yintercept = 3.24,linetype="dashed") +
  geom_hline(yintercept = 3,linetype="dashed",color="darkgray") +
  geom_point(size = 5,position = position_dodge(width = 0.5)) +
  scale_color_manual(values=c("black","darkgrey")) +
  scale_shape_manual(values=c(21,19)) +
  scale_fill_manual(values = c("white","white")) +
  scale_y_continuous(limits = c(0,10),breaks = seq(0,10,by=5)) +
  facet_grid(.~factor(micro,levels = c("Apo","LQ","HQ"))) +
  ylab(expression(atop("Biomass [g]","mean \u00B1 CI (n = 5)"))) +
  xlab("Phage addition") +
  theme_bw(base_size = 20) +
  theme(panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black",fill=NA),
        panel.background = element_rect(fill="white"),
        axis.text.x=element_blank(),
        axis.text.y=element_text(size=rel(0.9)),
        axis.title.x =element_blank(),
        axis.title.y = element_text(size=rel(0.9)),
        legend.position = "none")
```

#join three plots together
```{r}
flower <- flower + theme(plot.margin = margin(0.75, 0.75, 0.75, 0.75, "cm"))
stolon <- stolon + theme(plot.margin = margin(0.75, 0.75, 0.75, 0.75, "cm"))
mass <- mass + theme(plot.margin = margin(0.75, 0.75, 0.75, 0.75, "cm"))

legend <- get_legend( mass + 
                          theme(legend.position = "right", 
                                legend.justification = "center",
                                legend.box.just = "right")
            )

plot_grid(
  plot_grid(mass, flower, stolon, ncol = 1, nrow = 3, 
            labels = c("A.)", "B.)", "C.)"), 
            label_size = 20),
  legend,
  rel_widths = c(2, 0.2),  # Adjust the relative width if needed
  ncol = 2, 
  align = "v", 
  axis = "tb"
)
```
